hbm_paper_results

### 2024-03-20 paper_lbcc_results.Rmd
parsing through results, fix na.omit

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site_predictions.csv"))
dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site_predictions.csv"))
dim(gamlss_centile_df)
```

@@

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.site_pred_deltas.csv"))
# skimr::skim(deltas_df)
cent_deltas_df <- deltas_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

@@

sample_boxplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=interaction(site,study, sep = "&"), x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text=element_blank(), ggh4x.axis.nesttext.y = element_text(size=6)) +
  labs(color="Sex", y = "Site")

@@

sample_barplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site,study, sep = "&"), fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(color="Sex")

@@

cent_deltas_df_long %>%
  ggplot() +
  geom_histogram(aes(x=delta, color=study, fill=study))
```

```{r}
delta_n.sex <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(sex, study, site) %>%
  summarise(sex_n = n(),
            mean_abs.delta=mean(abs.delta),
            mean_delta=mean(delta),
            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each site

#get site-level stats
delta_summary.df <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study, site) %>%
  summarise(site_mean_abs.delta=mean(abs.delta),
            site_mean_delta=mean(delta),
            site_mean_age=mean(age_days),
            site_mean_age_yrs=mean(age_days)/365.25,
            site_max_abs.delta=max(abs.delta),
            site_n=n()) %>%
  mutate(pheno_cat="all")

# #get w/in cat stats
# delta_summary.cat.df <- cent_deltas_df_long %>%
#   mutate(abs.delta = abs(delta),
#          pheno_cat = case_when(
#     pheno %in% vol_list_global ~ "Global Volume",
#     pheno %in% vol_list_regions ~ "Regional Volume",
#     pheno %in% sa_list ~ "Regional SA",
#     pheno %in% ct_list ~ "Regional CT",
#     TRUE ~ NA_character_)) %>%
#   group_by(study, site, pheno_cat) %>%
#   summarise(site_mean_abs.delta=mean(abs.delta),
#             site_mean_delta=mean(delta),
#             site_mean_age=mean(age_days),
#             site_mean_age_yrs=mean(age_days)/365.25,
#             site_max_abs.delta=max(abs.delta),
#             site_n=n())
# 
# delta_summary.df <- full_join(delta_summary.df, delta_summary.cat.df)

delta_n.sex %>%
  dplyr::filter(site=="abide2_BNI_1")
```

```{r}
delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
  mutate(sex_prop=sex_n/site_n,
         majority = ifelse(sex_prop > .5, 1, 0))

delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
  pivot_wider(id_cols=c(study, site, site_n, site_mean_age, site_mean_age_yrs, site_mean_delta, site_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))

#Merge back female prop back into overall site-level stats
to_merge <- delta_sum.by.sex.df_piv %>%
  dplyr::select(study, site, sex_prop_Female) %>%
  mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study

@@

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=site, fill=site)) +
  geom_count(aes(size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site)) +
  theme(legend.position="none")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site), color="darkgray")

@@

delta_sum.by.sex.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta, color=study)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")
```

look at max abs. delta across sites
```{r}
delta_sum.df_merged %>%
  ggplot() +
  geom_histogram(aes(x=site_max_abs.delta*100))

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=site_n)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=site_mean_age_yrs)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
  geom_point() +
  geom_smooth(method="lm")
```
#Site Effects

Test to see if centile score is associated with site in either ComBatLS or ComBat-GAM harmonized data.

@@

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["site"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

@@

Test single pheno

```{r}
gam_centile_df.clean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=WMV)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

summary(lm(WMV ~ site + age_days + sex, data = gam_centile_df.clean))
```

@@

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

@@

```{r}
summary(lm(mean_cent ~ site, data=gam_centile_df.mean))
summary(lm(mean_cent ~ study, data=gam_centile_df.mean))

summary(aov(mean_cent ~ site, data=gam_centile_df.mean))
```

success!

@@

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["site"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))
```

@@

```{r}
summary(lm(mean_z ~ site, data=gam_z_df.mean))
summary(lm(mean_z ~ study, data=gam_z_df.mean))

summary(aov(mean_z ~ study, data=gam_z_df.mean))
```

@@

```{r}
gamlss_centile_df.clean %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=WMV))

wmv_cent_gamlss.lm <- lm(WMV ~ study + age_days + sex, data=gamlss_centile_df.clean)

# gtsummary::tbl_regression(wmv_cent_gamlss.lm) %>%
#   gtsummary::add_global_p()

gamlss_centile_df.clean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=WMV, fill=study)) +
  theme(axis.text.x = element_text(angle=90), legend.position = "none")
```

@@

df_gamlss_feat_aov <- data.frame()
for(pheno in feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["site"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

@@

df_gamlss_feat_aov.study %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  nrow()
```
average w/in subject

@@

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(site, age_days), y=mean_cent)) + 
  theme(axis.text.x = element_text(angle=90))

gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  theme(axis.text.x = element_text(angle=90))

@@

gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=mean_cent, fill=study)) + 
  theme(axis.text.x = element_text(angle=90), legend.position = "none")
```

@@

```{r}
summary(lm(mean_cent ~ site, data=gamlss_centile_df.mean))
summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))

summary(aov(mean_cent ~ site, data=gamlss_centile_df.mean))
```

success!

@@

```{r}
gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["site"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))
```

@@

```{r}
gamlss_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
  facet_wrap(~study)
```

@@

```{r}
summary(lm(mean_z ~ site, data=gamlss_z_df.mean))
summary(lm(mean_z ~ study, data=gamlss_z_df.mean))

summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
```

#Site-Level Mean Results

@@

#overall means
gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
  group_by(study, site) %>%
  summarize(mean_gam_cent=mean(centile)) %>%
  mutate(pheno_cat = "all")

@@

summary(lm(mean_gam_cent ~ site_mean_age, data=gam_centile_site.mean.df2))
summary(lm(mean_gam_cent ~ site_mean_age + sex_prop_Female + site_n, data=gam_centile_site.mean.df2))
```

@@

#overall means
gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
  group_by(study, site) %>%
  summarize(mean_gamlss_cent=mean(centile)) %>%
  mutate(pheno_cat = "all")

@@

summary(lm(mean_gamlss_cent ~ site_mean_age, data=gamlss_centile_site.mean.df2))
summary(lm(mean_gamlss_cent ~ site_mean_age + sex_prop_Female + site_n, data=gamlss_centile_site.mean.df2))
```

@@

Site-level mean centiles are not associated with study age in either ComBat-GAM or ComBatLS

@@

```{r}
# age
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))

#forward selection for site_mean_abs.delta
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.delta.age.fit <- lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space

full.delta.age.fit2 <- lm(site_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
summary(full.delta.age.fit2)
```

@@

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
  theme(axis.title.x=element_b)

>>>>

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
gam_centile_df <- gam_centile_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
gamlss_centile_df <- gamlss_centile_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(gamlss_centile_df)
```

@@

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
# skimr::skim(deltas_df)
cent_deltas_df <- deltas_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

@@

sample_boxplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=study, x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
  labs(color="Sex", y = "Study")

@@

sample_barplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(color="Sex")

@@

cent_deltas_df_long %>%
  ggplot() +
  geom_histogram(aes(x=delta, color=study, fill=study)) +
  theme(legend.position="none")
```

```{r}
delta_n.sex <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(sex, study) %>%
  summarise(sex_n = n(),
            mean_abs.delta=mean(abs.delta),
            mean_delta=mean(delta),
            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each site

#get study-level stats
delta_summary.df <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study) %>%
  summarise(study_mean_abs.delta=mean(abs.delta),
            study_mean_delta=mean(delta),
            study_mean_age=mean(age_days),
            study_mean_age_yrs=mean(age_days)/365.25,
            study_max_abs.delta=max(abs.delta),
            study_n=n()) %>%
  mutate(pheno_cat="all")
```

```{r}
delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
  mutate(sex_prop=sex_n/study_n,
         majority = ifelse(sex_prop > .5, 1, 0))

delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
  pivot_wider(id_cols=c(study, study_n, study_mean_age, study_mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))

#Merge back female prop back into overall site-level stats
to_merge <- delta_sum.by.sex.df_piv %>%
  dplyr::select(study, sex_prop_Female) %>%
  mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study

@@

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=study_mean_delta)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
  geom_count(aes(size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study)) +
  theme(legend.position="none")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study), color="darkgray")

@@

delta_sum.by.sex.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta)) +
  geom_count(aes(size=study_n, color=study)) +
  geom_smooth(method="lm")
```

look at max abs. delta across sites
```{r}
delta_sum.df_merged %>%
  ggplot() +
  geom_histogram(aes(x=study_max_abs.delta*100))

delta_sum.df_merged %>%
  ggplot(aes(y=study_max_abs.delta*100, x=study_n)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=study_max_abs.delta*100, x=study_mean_age_yrs)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=study_max_abs.delta*100, x=sex_prop_Female)) +
  geom_point() +
  geom_smooth(method="lm")
```
#Batch Effects

Test to see if centile score is associated with study in either ComBatLS or ComBat-GAM harmonized data.

@@

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

@@

@@

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

@@

```{r}
summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
```

@@

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gam_feat_aov.z.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

@@

```{r}
summary(lm(mean_z ~ study, data=gam_z_df.mean))
summary(aov(mean_z ~ study, data=gam_z_df.mean))
```

@@

@@

df_gamlss_feat_aov <- data.frame()
for(pheno in feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

@@

df_gamlss_feat_aov.study %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  nrow()
```

@@

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

@@

gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=mean_cent, fill=study)) + 
  theme(axis.text.x = element_text(angle=90), legend.position = "none")

gamlss_centile_df.mean %>%
  dplyr::filter(age_days>365) %>%
  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

@@

```{r}
summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
```

@@

```{r}
gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gamlss_feat_aov.z.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

@@

```{r}
gamlss_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
  facet_wrap(~study) +
  theme(legend.position="none")
```

@@

```{r}
summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
```

#Centile Fans
Look at brain charts for global vols

```{r}
gamlss.GMV.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_mod.rds"))

```

#Study-Level Mean Results

@@

#overall means
gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
  group_by(study) %>%
  summarize(mean_gam_cent=mean(centile),
            median_gam_cent=median(centile)) %>%
  mutate(pheno_cat = "all")

@@

summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
summary(lm(mean_gam_cent ~ study_mean_age + sex_prop_Female + study_n, data=gam_centile_site.mean.df2))

summary(lm(median_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
```

```{r}
#mean in global pheno
gam_centile_site.mean.df.glob <- gam_centile_df.clean_long %>%
  dplyr::filter(pheno %in% vol_list_global) %>%
  group_by(study) %>%
  summarize(mean_gam_cent=mean(centile),
            median_gam_cent=median(centile)) %>%
  mutate(pheno_cat = "global_vol")

gam_centile_site.mean.df.glob2 <- left_join(gam_centile_site.mean.df.glob, delta_sum.df_merged, by="study")

summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df.glob2))
```

@@

#overall means
gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
  group_by(study) %>%
  summarize(mean_gamlss_cent=mean(centile),
            median_gamlss_cent=median(centile)) %>%
  mutate(pheno_cat = "all")

@@

summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
summary(lm(mean_gamlss_cent ~ study_mean_age + sex_prop_Female + study_n, data=gamlss_centile_site.mean.df2))

summary(lm(median_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
```

plots

```{r}
all_means.df <- full_join(gam_centile_site.mean.df2, gamlss_centile_site.mean.df2)

all_means.df.long <- all_means.df %>%
  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))

all_means.df.long %>%
  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
  geom_point() +
  geom_smooth(method="lm")
```

Significant

```{r}
#mean in global pheno
gamlss_centile_site.mean.df.glob <- gamlss_centile_df.clean_long %>%
  dplyr::filter(pheno %in% vol_list_global) %>%
  group_by(study) %>%
  summarize(mean_gamlss_cent=mean(centile),
            median_gamlss_cent=median(centile)) %>%
  mutate(pheno_cat = "global_vol")

gamlss_centile_site.mean.df.glob2 <- left_join(gamlss_centile_site.mean.df.glob, delta_sum.df_merged, by="study")

summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df.glob2))

summary(lm(mean_gamlss_cent ~ study_mean_age + study_n + sex_prop_Female, data=gamlss_centile_site.mean.df.glob2))
```

@@

```{r}
all_means.df.glob <- full_join(gam_centile_site.mean.df.glob2, gamlss_centile_site.mean.df.glob2)

all_means.df.long.glob <- all_means.df.glob %>%
  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))

all_means.df.long.glob %>%
  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
  geom_point() +
  geom_smooth(method="lm")
```

@@

```{r}
# age
summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged))

#forward selection for site_mean_abs.delta
summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.delta.age.fit <- lm(study_mean_abs.delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
summary(full.delta.age.fit)

#forward selection for site_mean_delta
summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.age.fit <- lm(study_mean_delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
summary(full.age.fit)
```

test w/o babies < 2
```{r}
delta_sum.df_merged_noinf <- delta_sum.df_merged %>%
  dplyr::filter(study_mean_age_yrs >= 2)

nrow(delta_sum.df_merged) - nrow(delta_sum.df_merged_noinf)

summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))
summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))

all_means.df.long %>%
  dplyr::filter(study_mean_age_yrs >= 2) %>%
  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
  geom_point() +
  geom_smooth(method="lm")
```

@@

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
  theme(axis.title.x=element_blank())
